\section{Alternate conclusion}

The "information ecosystem" metaphor is a powerful way to understand complexes of data, people, and machines in a rapidly changing social and technological environment. At first glance it may suggest an interest in nature because of the presence of the eco- prefix and its common use as a term in the biological and environmental sciences. There is also a holistic implication that an ecosystem encompasses or contains the natural functions of a system. It is at this point when the concept of nature arises in our thoughts that the usage of the term ecosystem can become confused. The suggested presence of nature makes an ecosystem into a natural object, and this objectification makes the connection between a ecosystem and nature seem inevitable and given. 

When applied to human built systems, such as an economy, this implication of naturalism imbued in the very term being used begins to cloud and confuse the actual system being discussed. The problem with this confusion is that the very reason for using the term ecosystem in the first place, in order to emphasize the importance of nature and the non-human environment, is subtly elided and eliminated from the conversation. Nature disappears and we are left with  a human-built system, many times an economy, which may pay no more attention to the impacts of human activity on the environment as any other system we might have wished to condemn.

Consider the use of information ecology in the work of Davenport and Prusak. DP approach the issue as business management scholars who are trying to improve the human management processes for information within business organizations. From the beginning this is a human problem which only connects to nature at the most banal level of saying that all humans are part of nature. Human beings are part of nature, but that does not mean that everything human beings do is natural. This contrast is very difficult to explain clearly and applying the term ecosystem or ecology to human organizations and systems adds to the confusion.

Value judgments are made by the very use of the term - nature. To be natural is to be spontaneous, without processing or bias. But human organizations are artificial in the most basic sense - they are produced by human beings and maintained by human ingenuity and skill. They do not arise spontaneously, their organization is determined by rules, behavior, and hierarchies of human actions. In a sense they are fundamentally non-natural. To call these systems artificial is not derogatory, or at least it should not be consider that way. We should take pride in the artificial organizations we have created because they have accomplished some of the most complex tasks human beings are capable of, from forming governments to sending probes into outer space.

Acknowledging that these systems are artificial does not mean that the systems are in any way separated from nature. If ecology has taught us anything it is that human beings are deeply embedded in the natural world and cannot escape. Our organizations, systems, the vast quantities of information we have collected are artificial because we have built them. They are much more like autopoetic machines, like a complicated steam engine with a speed governor to limit the chances of explosion, than an organism imbued with life that adopts both cooperative and competetive strategies to reproduce. This incredible system of communication and data sharing is built inside of an ecosystem nevertheless, in the most accurate use of the term, that includes human beings, all other forms of life and the non-living materials upon which we rely to reproduce. Instead of referring to every collection of parts found in human civilization as an ecosystem we can simply refer to what they are, artificial systems or economies which have been built by human beings for a specific purpose.

Prepending the word information to ecosystems does nothing to clarify the matter because the word \textit{information} is almost as hard to define as \textit{nature}. But there are many other uses of the term ecosystem that can be abandoned more quickly. When a newspaper story refers to the “music ecosystem” we should just prescribe that usage as laziness. What is really being referred to is the music economy or the music business, the complex of artificial organizations, individuals, and technology which human beings have built to distribute music.

When applying the term ecology or ecosystem to technology and information the same laziness is never far away. Phrases like “big data ecosystem” or “open source ecosystem” are more accurately referring to the economic, business, and technological structures which support these human activities. As the title of the paper suggests, information has long been recognized as a part of the ecosystem. Information broadly conceived is what mediates the relationships between living organisms and their non-living environment. While metaphors may be as useful way to understand novel phenomena, their forms and roles within learning processes are still under debate within the philosophy of science literature. Like Bohr's planetary model of the atom, some metaphors are destined for retirement as our understandings of the phenomena deepen. Perhaps it is time to retire the information ecosystem metaphor as we begin to better understand the data revolution.

 In the case of human beings, information that we gather from our surrounding environment is what we use to make decisions about how we govern and use natural resources. Information has always been and will always be a part of this ecosystem, but to conceive of an information ecosystem that is separate from our relationship with nature may not be in our best interest as but one species that makes a home on planet earth. While metaphors may be as useful way to understand novel phenomena in scientific settings, their forms and roles within learning processes are still under debate within the philosophy of science literature. Like Bohr's planetary model of the atom, some metaphors are destined for retirement as our understandings of the phenomena deepen. Perhaps it is time to retire the information ecosystem metaphor as we begin to better understand the data revolution.